---
title: "{mscsweblm4r} vignette"
author: "Phil Ferriere"
date: "May 2016"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{mscsweblm4r vignette}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---
```{r echo=FALSE}
# Build with devtools::install(build_vignettes = TRUE)
library("knitr")
knitr::opts_chunk$set(comment = "#>", collapse = TRUE)
```

## Package Installation

You can either install the latest **stable** version from CRAN:

```{r eval=FALSE}
if ("mscsweblm4r" %in% installed.packages()[,"Package"] == FALSE) {
  install.packages("mscsweblm4r", repos = "http://cran.us.r-project.org")
}
```

Or, you can install the **development** version

```{r eval=FALSE}
if ("mscsweblm4r" %in% installed.packages()[,"Package"] == FALSE) {
  if ("devtools" %in% installed.packages()[,"Package"] == FALSE) {
    install.packages("devtools", repos = "http://cran.us.r-project.org")
  }
  devtools::install_github("philferriere/mscsweblm4r")
}
```

## Package Configuration

**This must be done BEFORE the package can load!**

At library attach time, `{mscsweblm4r}` will first check to see if the variable
`MSCS_WEBLANGUAGEMODEL_CONFIG_FILE` exists in the system environemt. If it does,
the package will use that as the path to the configuration file.

If `MSCS_WEBLANGUAGEMODEL_CONFIG_FILE` doesn't exist, it will look for the file
`.mscskeys.json` in the current user's home directory (that's `~/.mscskeys.json`
on Linux, and something like `C:\Users\Phil\Documents\.mscskeys.json` on
Windows). If the file is found, the package will load the API key and URL from
it.

If using a file, please make sure it has the following structure:

```json
{
  "weblanguagemodelurl": "https://api.projectoxford.ai/text/weblm/v1.0/",
  "weblanguagemodelkey": "...MSCS Web Language Model API key goes here..."
}
```

If no configuration file was found, `{mscsweblm4r}` will attempt to pick up its
configuration from two Sys env variables instead:

`MSCS_WEBLANGUAGEMODEL_URL` - the URL for the Web LM REST API.

`MSCS_WEBLANGUAGEMODEL_KEY` -  your personal Web LM REST API key.

## Error Handling Not Optional

The MSCS Web LM API is a **[RESTful](https://en.wikipedia.org/wiki/Representational_state_transfer)** API. HTTP requests over a network and the
Internet can fail. Because of congestion, because the web site is down for
maintenance, because of firewwall configuration issues, etc. There are many
possible points of failure.

The API can also fail if you've **exhausted your call volume quota** or are **exceeding
the API calls rate limit**. Unfortunately, MSCS does not expose an API you can query to check
if you're about to exceed your quota for instance. The only way you'll know for
sure is by **looking at the error code** returned after an API call has failed.

Therefore, you must write your R code with failure in mind. Our preffered way is
to use `tryCatch()`. Its mechanism may appear a bit daunting at first, but it
is well [documented](http://www.inside-r.org/r-doc/base/signalCondition). We've
also included many examples, as you'll see below.

## Package Loading with Error Handling

Here's some sample code that illustrates how to use `tryCatch()`:

```{r}
tryCatch({

  library('mscsweblm4r')

}, error = function(err) {

  geterrmessage()

})
```

If `{mscsweblm4r}` cannot locate `.mscskeys.json` nor any of the configuration
environment variables, the code above will generate the following output:

```{r eval=FALSE}
Error : .onAttach failed in attachNamespace() for 'mscsweblm4r', details:
  call: NULL
  error: mscsweblm4r: could not find MSCS_WEBLANGUAGEMODEL_KEY or MSCS_WEBLANGUAGEMODEL_URL in Sys env nor locate ~/.mscskeys.json
[1] "package or namespace load failed for 'mscsweblm4r'"
```

Similarly, package load will fail if `{mscsweblm4r}` cannot find the
`weblanguagemodelkey` key in `.mscskeys.json`, or fails to parse it correctly, etc.
This is why it is so important to use `tryCatch()`.

## Package API

The five API calls exposed by `{mscsweblm4r}` are the following:

```{r eval = FALSE}
  # Retrieve a list of supported web language models
  weblmListAvailableModels()
```

```{r eval = FALSE}
  # Break a string of concatenated words into individual words
  weblmBreakIntoWords(
    textToBreak,                      # ASCII only
    modelToUse = "body",              # "title"|"anchor"|"query"(default)|"body"
    orderOfNgram = 5L,                # 1L|2L|3L|4L|5L(default)
    maxNumOfCandidatesReturned = 5L   # Default: 5L
  )
```

```{r eval = FALSE}
  # Get the words most likely to follow a sequence of words
  weblmGenerateNextWords(
    precedingWords,                  # ASCII only
    modelToUse = "title",            # "title"|"anchor"|"query"(default)|"body"
    orderOfNgram = 4L,               # 1L|2L|3L|4L|5L(default)
    maxNumOfCandidatesReturned = 5L  # Default: 5L
  )
```

```{r eval = FALSE}
  # Calculate joint probability a particular sequence of words will appear together
  weblmCalculateJointProbability(
    inputWords =,                    # ASCII only
    modelToUse = "query",            # "title"|"anchor"|"query"(default)|"body"
    orderOfNgram = 4L                # 1L|2L|3L|4L|5L(default)
  )
```

```{r eval = FALSE}
  # Calculate conditional probability a particular word will follow a given sequence of words
  weblmCalculateConditionalProbability(
    precedingWords,                  # ASCII only
    continuations,                   # ASCII only
    modelToUse = "title",            # "title"|"anchor"|"query"(default)|"body"
    orderOfNgram = 4L                # 1L|2L|3L|4L|5L(default)
  )
```

These functions return S3 class objects of the class `weblm`. The `weblm` object
exposes formatted results (in `data.frame` format), the REST API JSON response
(should you care), and the HTTP request (mostly for debugging purposes).

## Sample Code

The following code snippets illustrate how to use {mscsweblm4r} functions and
show what results they return with toy examples. If after reviewing this code
there is still confusion regarding how and when to use each function, please
refer to the [original documentation](https://www.microsoft.com/cognitive-services/en-us/web-language-model-api/documentation).

### List Available Models function

```{r}
tryCatch({

  # Retrieve a list of supported web language models
  weblmListAvailableModels()

}, error = function(err) {

 # Print error
 geterrmessage()

})
```

### Break Into Words function

```{r}
tryCatch({

  # Break a sentence into words
  weblmBreakIntoWords(
    textToBreak = "testforwordbreak", # ASCII only
    modelToUse = "body",              # "title"|"anchor"|"query"(default)|"body"
    orderOfNgram = 5L,                # 1L|2L|3L|4L|5L(default)
    maxNumOfCandidatesReturned = 5L   # Default: 5L
  )

}, error = function(err) {

  # Print error
  geterrmessage()

})
```

### Generate Next Word function

```{r}
tryCatch({

  # Generate next words
  weblmGenerateNextWords(
    precedingWords = "how are you",  # ASCII only
    modelToUse = "title",            # "title"|"anchor"|"query"(default)|"body"
    orderOfNgram = 4L,               # 1L|2L|3L|4L|5L(default)
    maxNumOfCandidatesReturned = 5L  # Default: 5L
  )

}, error = function(err) {

  # Print error
  geterrmessage()

})
```

### Calculate Joint Probability function

```{r}
tryCatch({

  # Calculate joint probability a particular sequence of words will appear together
  weblmCalculateJointProbability(
    inputWords = c("where", "is", "San", "Francisco", "where is",
                   "San Francisco", "where is San Francisco"),  # ASCII only
    modelToUse = "query",                     # "title"|"anchor"|"query"(default)|"body"
    orderOfNgram = 4L                         # 1L|2L|3L|4L|5L(default)
  )

}, error = function(err) {

  # Print error
  geterrmessage()

})
```

### Calculate Conditional Probability function

```{r}
tryCatch({

  # Calculate conditional probability a particular word will follow a given sequence of words
  weblmCalculateConditionalProbability(
    precedingWords = "hello world wide",       # ASCII only
    continuations = c("web", "range", "open"), # ASCII only
    modelToUse = "title",                      # "title"|"anchor"|"query"(default)|"body"
    orderOfNgram = 4L                          # 1L|2L|3L|4L|5L(default)
  )

}, error = function(err) {

  # Print error
  geterrmessage()

})
```
